---
title: "Forecasting Demand v.01"
author: "Arif P. Sulistiono / @arifpras"
date: "2022-11-10"
output: html_document
---

```{r prerequisites, include=FALSE}
# From {usethis} to avoid loading a massive .RData file
usethis::use_blank_slate(scope = c("user", "project"))

# Removes all items in Environment!
rm(list = ls())
ls()

# Load packages
pacman::p_load(tidyverse, seasonal, fpp2, readxl, feasts, devtools, tsibble)

# My ggplot's themes
devtools::source_url("https://raw.githubusercontent.com/arifpras/themesAPS/main/themesAPS?raw=TRUE")

options(tz = "Asia/Jakarta")

# https://www.kaggle.com/code/brwalk/an-introduction-to-time-series-forecasting-in-r
# https://www.mitchelloharawild.com/blog/feasts/
```

```{r working-dir, include=FALSE}
setwd("~/Dropbox/01_Kemenkeu/2022_11/forecastingDemand")
```

```{r load}
db01 <-
  read_excel(
    "20221108_database.xlsx",
    sheet = "db00",
    col_types = c("date", "numeric", "numeric", "numeric", "numeric")
  ) %>% 
  select(-bid_10e6, -dpk_10e12, -vix_index) %>% 
  glimpse()
```

```{r time_series}
(db02 <- ts(db01$bid_10e12))
```

```{r update_ts}
# convert as time series
(db03 <- ts(db01$bid_10e12, frequency = 12, start = 2016))

# convert as tsibble
(db03_tsbl <- as_tsibble(db03, key = bid_10e12, index = month))
```

# Forecasting Method 1: Exponential Smoothing

```{r autoplot}
(plot01 <- autoplot(db03, main = "Government Debt Securities' demand: Jan. 2016-Oct. 2022") + theme_light() + theme_aps02)

# tsibble plot
(plot01_tsbl <-
    db03_tsbl %>% autoplot(value) + theme_light() + theme_aps02)
```

```{r seasonal_plot}
(plot02 <- ggseasonplot(db03, main = "Seasonal Plot: Government Debt Securities' demand") + theme_light() + theme_aps02)

#tsibble
# (db03_tsbl %>% gg_season(value) + theme_light() + theme_aps02)
db03_tsbl %>% gg_subseries(value) + theme_light() + theme_aps02
```

```{r additive}
db04a <- decomposed_sales_additive <- decompose(db03, type = "additive")

(plot03a <- autoplot(db04a)+ theme_light() + theme_aps02)
```

```{r multiplicative}
db04b <- decomposed_sales_additive <- decompose(db03, type = "multiplicative")

(plot03b <- autoplot(db04b)+ theme_light() + theme_aps02)
```

Types of seasonal decompositions

-   Additive Seasonal Decomposition: Each season gets moved by a constant number that is added or subtracted from the trend (e.g. subtract \$100 to sales of swimsuits in December)

-   Multiplicative Seasonal Decomposition: Each season has a number we multiply to the trend (e.g. We multiply swimsuit sales by 1.2 in July)

If we know that seasonal affects do not change over time an additive, then an additive decomposition is great. When the seasonal affects grow with the trend, a multiplicative decomposition is better. If you notice the residuals in your additive model are very large when the trend is at a low or high, you may want to switch to a multiplicative model.

Why is the additive selected? Because usually there is no issuance in December, so the trend needs to be subtracted.

```{r ses_method}
(plot04 <- autoplot(ses(db03), PI = FALSE) + theme_light() + theme_aps02)
```

```{r holt_method}
(plot05 <-
   (autoplot(holt(db03), PI = FALSE) + theme_light() + theme_aps02))

plot05_build <- ggplot_build(plot05) # extracting the data from plot05, sublists: data [1] raw; [2] forecasting
# layer_data(plot05, 1)
plot05_build$data
```

```{r holt_dampened}
(plot06 <-
   autoplot(holt(db03, damped = TRUE, h = 14), PI = FALSE) + theme_light() + theme_aps02)
```

```{r guererro}
db03lambda <-
  db03_tsbl %>% features(value, features = "guererro") %>% 
  pull(lambda_guerrero)
```

```{r ets_method}
(plot07 <-
   autoplot(forecast(ets(db03), h = 14, PI = FALSE)) + 
   theme_light() + theme_aps02) 
```

```{r decompose_ets}
(plot08 <- autoplot(ets(db03)) + theme_light() + theme_aps02)
```

```{r summary_ets}
summary(ets(db03))
```

# Forecasting Method 2: Autoregression

```{r autoreg_plot}
(plot09 <- ggAcf(db03) + theme_light() + theme_aps02)
```

```{r autoreg_db03}
autoregression_db03 <- ar(db03)
autoregression_db03
```

The ar() has recommended we use an order of 12 Meaning we use the previous 12 values to predict the next one. The coefficients are also shown to us.

```{r ar_forecast}
(plot10 <-
   autoplot(forecast(
     autoregression_db03, level = 0, h = 14
   )) + theme_light() + theme_aps02)

# level = 0 prevents a confidence interval from being generated
```

```{r decomposed_db03}
decomposed_db03 <- seas(db03) 
# time series decomposition

decomposed_db03 <-
  data.frame(decomposed_db03$data) 
# makes a data frame out of the output of seas() which is not a data frame

time_series_db03_trend <-
  ts(decomposed_db03$trend,
     frequency = 12,
     start = 2016) 
# turn the trend into a time series

ar_forecast <-
  forecast(ar(time_series_db03_trend), h = 14) 
# forecast the trend
```

```{r ar_forecast_plot}
(
  plot11 <-
    autoplot(db03) +
    autolayer(ar_forecast$mean * tail(decomposed_db03$seasonal, 14),
              color = "blue") +
    theme_light() + theme_aps02
)
# ar_forecast$mean gives us our forecast

# We multiply by the last 12 seasonal factors to resesonalize, using tail
# Use tail() to select the last values of something, passing in 12 to give us the most recent seasonal factors. tail() works like head()

# autolayer() adds another series to a plot
```

# Forecasting Method 3: ARIMA

```{r arima01}
(plot12 <- autoplot(decompose(db03)) + theme_light() + theme_aps02)
```

```{r arima02}
(plot13 <-
   autoplot(forecast(Arima(db03, order = c(
     1, 1, 1
   )), h = 14), PI = FALSE) + theme_light() + theme_aps02)
```

```{r arima_plot}
(plot14 <-
   autoplot(forecast(auto.arima(db03), h = 14), PI = FALSE) + theme_light() + theme_aps02)
```

# Forecasting Method 4: Regression

```{r db_regress}
db05 <-
  read_excel(
    "20221108_database.xlsx",
    sheet = "db00",
    col_types = c("date", "numeric", "numeric", "numeric", "numeric")
  ) %>% 
  select(-bid_10e6) %>% 
  mutate(
    month = as.Date(month, "%Y-%m-%d")
  ) %>% 
  glimpse()
```

```{r ts_regress}
lm01 <- lm(bid_10e12 ~ lag(dpk_10e12, n = 1L), data = db05)

summary(lm01)
```

```{r cleaning, include=FALSE}
# Removes all items in Environment!
rm(list = ls())
ls()
```
